\documentclass{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{xcolor}

\usepackage{listings}
\usepackage{mdframed}
\usepackage{fourier}

\lstset{
	basicstyle=\small,
	keywordstyle=\color{black}\bfseries
}

\usepackage{enumerate}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=True]{hyperref}

\title{Notes \\ Seminar on Machine Learning and Neural Networks}
\date{}
\author{}

\DeclareMathOperator{\Inp}{Inp}
\DeclareMathOperator{\Tr}{Tr}

\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}

\begin{document}

\maketitle

% \chapter*{Preface}

% This notes are the result of a seminar that took place on the first semester of 12018 at the National University of Colombia at Medellín. The objective of the seminar was to cover the basics of Machine Learning needed to understand neural networks. To do so, we followed a rather theoretical approach to the subject. Our main references were \cite{bishop1} and \cite{bishop2} (and these books are so well made that most of our notes are almost verbatum copies from them). For a holistic understanding of the subject, though, we implemented most examples using python and R and following a [MOOC] on Udemy.

\include{chapter1_linear_regression}

\chapter{Linear Classification}

Classification problems consist on constructing a model that takes every $x$ in the input space and maps it to a discrete set $\{C_1,\dots, C_K\}$ of \textit{classes}, given that we already have a set of classified data $\mathcal{D} = \{(x_i, t_i)\}_{i=1}^N$. There are plenty of examples of real life problems that require a classification, for example
\begin{itemize}
	\item We have diagnostic images of a liver, and we want to determine wheter a new patient has liver cancer or not.

	\item We have a collection of labeled handwritten digits (such as \href{http://yann.lecun.com/exdb/mnist/}{the MNIST database}), and we want to determine which digit appears in a new image.

	\item We have a song, and we want to determine whether it is a happy, upbeat one or a melancholic one.

	\item A bank may have plenty of information about their clients, and may want to construct a model to determine if a new client will default on their credit or not.

	\item \href{https://cs.stanford.edu/people/esteva/nature/}{A Nature article} explains how a model for classifying skin lesions at dermatologist-level was constructed.
\end{itemize}

Many of these examples may require heavy machinery, but for now we will introduce the models of classification by considering the most basic ones: linear models of classification.

There are three ways to construct a model, and they go (in increasing level of complexity) like this:

\begin{enumerate}[(a)]
	\item Model a \textbf{discriminant function} $y(x,w)$ that classifies $x$.

	\item Model $P(C_k|x)$ for $k\in\{1,\dots,K\}$ and use this information to take decisions over $x$, such as classifying $x$ to the class $C_k$ such that $P(C_k|x) \geq P(C_j|x)$ for all $j\in\{1,\dots, K\}$. Such models are called \textbf{Probabilistic Discriminant Functions}, because the $y_k(x) = P(C_k|x)$ now serve as discriminant functions.

	\item Model $P(x|C_k)$ and $P(C_k)$ (i.e. modeling the joint distribution $P(x, C_k)$) and then use Bayes to get $P(C_k|x)$. These models are called \textbf{Generative Models}, because we can marginalize the joint distribution for the $x$ and generate new training data.
\end{enumerate}

\section{Linear Discriminant Models}

Our first approach to classification will be using linear discriminant models, we will learn the parameters of a series of functions of the form

\[y_k(x,w_k) = w_k^Tx + w_{k0},\, k\in\{1,\dots, K\}\]

such that we classify $x$ to $C_k$ if $y_k(x,w_k) > y_j(x,w_j)$ for all $j\neq k$. Naming names, we call $w_k = (w_{k1},\dots, w_{kK})^T$ the parameters, $w_{k0}$ is a special parameter called the \textbf{bias}, and $-w_{k0}$ is called the \textbf{threshold}. We can get all these linear combinations together into a vector and consider only one discriminant function

\[y(x,w) = \begin{bmatrix}
	y_1(x,w_1)\\
	y_2(x,w_2)\\
	\vdots\\
	y_K(x,w_K)
\end{bmatrix} = W^Tx + W_0\]

where $W$ is a matrix whose columns are precisely $w_1, w_2,\dots, w_K$, and $W_0 = (w_{1,0}, w_{2,0},\dots, w_{K,0})^T$. Such a model is usually called a 1-of-$K$ scheme.

\subsection{The geometry of this model}

Let's consider the geometry that such a scheme induces. Call the regions $R_k := \{x\;|\;x \mbox{ is classified to } C_k\}\subseteq \Inp$ and $\partial R_k$ the \textbf{decision boundaries}. We claim that these decision boundaries are hyperplanes and that the $R_k$s are simply connected and convex.

Indeed, the first claim follows from the fact that the decision boundary between $R_i$ and $R_j$ is given by $\{x\;|\;y_i(x) = y_j(x)\}$, because then
\begin{align*}
y_i&(x) = y_j(x)\\
&\Rightarrow w_i^Tx + w_{i0} = w_j^Tx + w_{j0}\\
&\Rightarrow (w_i - w_j)^Tx + (w_{i0} - w_{j0}) = 0
\end{align*}
thus, the decision boundary is given by a hyperplane. Now let's prove that each region is convex and simply connected: let $x,x'\in R_k$, consider then a convex combination $\hat{x} = \lambda x + (1-\lambda)x'$, we need to prove that for every $\lambda\in [0,1]$, $\hat{x}\in R_k$, i.e., $y_k(\hat{x}) > y_j(\hat{x})$ for every $j\neq k$. Indeed,

\begin{align*}
 	y_k(\hat{x}) &= y_k(\lambda x + (1-\lambda)x')\\
 	&= w_k^T(\lambda x + (1-\lambda)x') + w_{k0}\\
 	&= \lambda(w_k^T x + w_{k0}) + (1-\lambda)(w_k^Tx' + w_{k0})\\
 	&= \lambda y_k(x) + (1-\lambda)y_k(x')\\
 	&> \lambda y_j(x) + (1-\lambda)y_j(x')\\
 	&= y_j(\hat{x})
 \end{align*}

(notice that we skipped the same algebraic process between the last steps). This proves that the model we're dealing with splits the regions into convex polytopes. This is a rigid geometry!, we're not to expect that the actual pattern behind the data be \textbf{linearly separable}. This is not generally the case, and we require more sophisticated methods in order to classify data that's not very linearly separable.

\subsection{Learning the parameters}

We now have a linear model, now let's talk about a couple ways we can \textit{learn} the parameters from training data.

\subsubsection{Sum-of-squares}

We can draw an analogy with regression, and try to fit a function using least squares. We could define an error associated with the training data $\mathcal{D} = \{(x_i, t_i)\}_{i=1}^N$:

\[E_\mathcal{D}(w_1,\dots,w_k) = \frac{1}{2}\sum_{k = 1}^K\sum_{n=1}^N(x_n^Tw_k + w_{k0} - t_{nk})^2\]

in order to get the minima of this error, we can rewrite it as the trace of a matrix. Let

\[\widetilde{X} = \begin{bmatrix}
	1, x_1^T\\
	% \hline
	\cdots\\
	% \hline
	1, x_N^T
\end{bmatrix}_{N\times(D+1)} = \begin{bmatrix}
	\widetilde{x}_1^T\\
	% \hline
	\cdots\\
	\widetilde{x}_N^T
\end{bmatrix}_{N\times(D+1)}\]
\[\widetilde{W} = \begin{bmatrix}
	\widetilde{w}_{1}\,|\,\cdots\,|\,\widetilde{w}_K
\end{bmatrix}_{(D+1)\times K}\]

where $\widetilde{w}_k = (w_{k0}, w_{k1}, \dots, w_{kK})^T = (w_{k0}, w_k^T)^T$, and finally

\[T = \begin{bmatrix}
	t_1^T\\
	\cdots\\
	t_N^T\\
\end{bmatrix}_{N\times K} \]

Then the error above defined can be written as

\[E_\mathcal{D}\left(\widetilde{W}\right) = \frac{1}{2}\Tr\left((\widetilde{X}\widetilde{W} - T)^T(\widetilde{X}\widetilde{W} - T)\right)\]

we can differentiate with respect to $\widetilde{W}$ in order to get the minimum, which is given by

\[\widetilde{W}^* = (\widetilde{X}^T\widetilde{X})^{-1}\widetilde{X}^TT\]

Notice that this is the same solution as we got in the linear regression scenario, and that the Moore-Penrose pseudoinverse pops up once more.

\subsubsection{The perceptron algorithm}

[to be filled]

\section{Probabilistic Discriminative Models}





\end{document}